Computing fusion products of MV cycles using the Mirkovi\'c--Vybornov isomorphism

### 2021-06-11 submission.tex
ready steady

%!TEX program = lualatex
% 
\documentclass[draft]{article} % [draft] TODO: use this at the end to check for overfills 
\usepackage{basic}
\usepackage{emoji}
\usepackage{float}
% TODO: The matrix is lower triangular and the lower-triangular matrix is
% \setemojifont{Apple Color Emoji}
% Working title: 
\title{How to compute the fusion product of MV cycles in type $A$}
\author{Roger Bai, Anne Dranowski, Joel Kamnitzer} 
% \date{October 2020}
\date{\today}

@@

\maketitle
\tableofcontents

\section{Todolist}
\begin{itemize}
    \item Remove the TOC 
    \item a shorthand: in examples, instead of writing $Z(\tau)$ for the MV cycle corresponding to the tableau $\tau$ we will only the tableau $\tau$!
    \item our $B(\infty) \to B(\lambda)$ does not match that of Claxton--Tingley or Hong--Lee but discussing how the two are related might be irrelevant since we don't worry about the crystal structure on MV cycles anyway (let alone investigate possible compatibility with the crystal structure on SSYT given $\tau\mapsto Z(\tau)$); \acom{perhaps we could mention that we don't know if this map is compatible with crystal structures} (but suspect there is a way of making it so); 
    \item Edit the intro 
    \item We are using the word [geometric] fusion just for the central fibre, we also have arithmetic fusion (aka multiplication in $\CC[N]$), and BD family: $X\geofu Y, b(X) \arfu b(Y), X \bdfamc Y\supset X\bdfamo Y$
    \item[\checkmark] Which MVy paper to cite? Both
    \item Super repetitive about lambdas and mus 
    \item Formatting could be bit better, e.g.\ some stuff goes outside the margin
    \item Fix section titles and other organizational seasoning  
    \item Fix numbering so that theorems, defs etc share a counter and are restarted by section 
    \item tableau of weight/filling? box of weight/filling? I prefer weight but don't mind filling 
    \item Recalling intersection multiplicity, checking that the definitions allow passing to open sets
    \item[\checkmark] in examples put ideals of arithmetic fusion and their tableaux into tables; also split examples into 2 sections: $\GL_3$ and $\GL_4$; replace $Z_M$ by $Z(\tau)$ and $Z(\tau)$ by $\tau$; and let's use $A$ for matrices instead of $X$ to be consistent with rising action?
\end{itemize}

\section{Introduction}
\label{s:intro}
% 
%Throughout this paper (unless indicated otherwise) the base field is $\CC$, but it does not have to be?  
\subsection{Geometric Satake and Mirkovic--Vilonen cycles}
The geometric Satake equivalence of Mirkovi\'c and Vilonen \cite{mirkovic2007geometric} provides an equivalence of categories between the category of spherical perverse sheaves in the affine Grassmannian of a reductive group $ G $ and the category of representations of its Langlands dual group $G^\vee$.  This foundational result provides a powerful tool to study representation theory using geometry.  In particular, under this equivalence, the Mirkovi\'c--Vilonen cycles in the affine Grassmannian of $ G $ index bases for irreducible representations of $ G^\vee $.  In this paper we work with $ G = \GL_m$ and identify $ G^\vee = \GL_m$ as well.

% The Mirkovi\'c--Vybornov isomorphism
In \cite{mirkovic2007quiver} (see also the recent sequel \cite{mirkovic2019comparison}) Mirkovi\'c and Vybornov provide a geometric version of symmetric and skew Howe $(\GL_m,\GL_N)$ dualities.  They relate Kazhdan--Lusztig slices in the affine Grassmannian of $\GL_m$ to slices in $N\times N$ nilpotent orbits on the one hand, 
% conjugacy classes of nilpotent matrices 
and to Nakajima quiver varieties on the other hand.    The second author \cite{dthesis} showed  that under the Mirkovi\'c-Vybornov isomorphism, the MV cycles are identified with certain varieties of matrices, called generalized orbital varieties.

%This extends to an isomorphism between families of slices in the Beilinson--Drinfeld and families of matrices with two eigenvalues. 
%This paper generalizes the first of these relations (an isomorphism) to families of slices in the Beilinson--Drinfeld Grassmannian. \jcom{I changed this because MVy did consider the BD Grassmannian as well.}

Essential to \cite{mirkovic2007geometric}, the Beilinson--Drinfeld Grassmannian is used to define a fusion product on the category of spherical perverse sheaves on the affine Grassmannian.  In \cite{anderson2003polytope}, Anderson used the Beilinson--Drinfeld Grassmannian to define a fusion of Mirkovi\'c--Vilonen cycles.  He constructed a family where the general fibre is a Cartesian product of two MV cycles, and the special fibre, called the fusion, is a (non-reduced) union of MV cycles.  Anderson conjectured that this fusion product matches the multiplication in the coordinate ring of the unipotent subgroup $ N \subset\GL_m$.  This conjecture was proven in \cite{baumann2019mirkovic}, where it was also established that the MV cycles give a basis for $ \CC[N]$.
% 
%To quote from \cite{mirkovic2007geometric}:
%\begin{quotation}
%    However, we make crucial use of an idea of Drinfeld, going back to around 1990. He discovered an elegant way of obtaining the commutativity constraint by interpreting the convolution product of sheaves as a ``fusion'' product.
%\end{quotation}
%

\subsection{Fusion using generalized orbital varieties}
% \acom{imho, fortune has nothing to do with it :)}
% Unfortunately, c
Computing the fusion product of MV cycles using the BD Grassmannian is quite difficult, both computationally and conceptually.  A few such computations were done by Anderson-Kogan in \cite{anderson2006algebra} in a somewhat ad-hoc fashion.  Another perspective on these computations (using convolution Grassmannians) was recently studied by Baumann, Gaussent \& Littelmann \cite{baumann2020bases}.

The purpose of this paper is to give a conceptually elementary way to compute this product by transferring it to a fusion product of generalized orbital varieties.  To do so, we use that the Mirkovic-Vybornov isomorphism extends to an isomorphism between families of slices in the Beilinson--Drinfeld and families of slices of matrices with two eigenvalues. Our main result is the following.

\begin{theorem}
    Let $ Z, Z' $ be MV cycles.  The structure constants for the multiplication $ b_Z b_{Z'}$ in $\CC[N]$ are equal to the intersection multiplicities in the fusion of the generalized orbital varieties corresponding to $ Z, Z'$.
\end{theorem}
%

@@

%One should care about the fusion product on bases \emoji{imp} \emoji{grin} \emoji{worried} \emoji{monocle-face} \emoji{skull} if one is interested in positivity!
\subsection{Cluster algebras and MV cycles}
One motivation for computing products of MV basis elements is to relate the cluster algebra structure on $\CC[N]$ to the MV basis. Cluster algebras were introduced by Fomin and Zelevinsky in \cite{fomin2002cluster1} and it was shown in \cite{geiss2007initial}, using the construction of \cite{berenstein2005cluster3}, that $\CC[N]$ has a natural cluster algebra structure. 
%\acom{As a reader I am confused - did both of them show it independently? Can I go to either source to find out how this works? I suggest ``in \cite[\S 4.3]{geiss2007initial} using \cite[Theorem~2.10]{berenstein2005cluster3}''} \rcom{I agree with the suggestion.}

This structure starts with an initial seed $\Sigma = ((x_1,\dots,x_r),B)$, where $x_i \in \CC[N]$, $B$ is a certain matrix, and $ r = m(m-1)/2$.
% ($=\#\Delta_+$)
We obtain new variables through a process called mutation where for a mutable variable $x_i$, its mutation is the unique element $x_i^* \in \CC[N]$ such that $$x_ix_i^* = x_+ + x_-.$$
This equation is called an exchange relation and the terms $x_+$ and $x_-$ are certain monomials in the $x_j$, $j\neq i$,  determined by $B$. 
%We then obtain a new seed $$\mu_i(\Sigma) = ((x_1,\cdots, x_{i-1}, x_i^*, x_{i+1},\cdots,x_r), \mu_i(B))$$ and can acquire possibly new variables through further mutations.

The variables $x_1,\cdots,x_r$ as well as all possible variables that are obtained through successive mutations are known as cluster variables. The cluster monomials are products of cluster variables that are supported on a single seed.

%\acom{It would be nice to add a line to set up this paragraph. Else the canonical bases come out of nowhere? Like ``There is another nice structure on $\CC[N]$, that of a \new{perfect basis} introduced by/studied by...; and heaps of work (or just the following two sources) suggests that the two structures are inextricably linked.'' Because it's a paper and not a talk.}
%\rcom{Something along these lines?}
There are two well-known bases of $\CC[N]$ whose relationship with the cluster structure has been studied heavily: the dual canonical basis \cite{lusztig1990canonicalbases} and the dual semicanonical basis \cite{lusztig2000semicanonical}.
Kang, Kashiwara, Kim \& Oh showed that all cluster monomials are contained in the dual canonical basis \cite{kang2018monoidal}, while Gei\ss, Leclerc, \& Schr\"oer proved that they are all contained in the dual semicanonical basis \cite[Theorem~2.8(3)]{geiss2006rigid}.
% 
%It has been shown that the cluster monomials are in certain interesting bases such as the dual canonical basis \cite{lusztig1990canonicalbases}\cite{kang2018monoidal} and the dual semicanonical basis \cite{lusztig2000semicanonical}\cite{geiss2006rigid}. As these elements are defined in a combinatorial way, it allows efficient computation of several elements of these bases.
% 
%\acom{Re: above paragraph, I actually dug up the precise references in rewriting this so maybe we can preserve that...}
%a geometric construction of a $\CC$-basis of $\cU\n$ the universal enveloping algebra of the Lie algebra of $N$, called the semicanonical basis. This becomes the dual semicanonical basis of the graded dual of $\cU\n$ which is isomorphic (as a Hopf algebra) to $\CC[N]$. 
%Geiss, Leclerc \& Schroer prove that certain cluster monomials are contained in Lusztig's (dual) semicanonical basis \cite[Theorem~2.8(3)]{geiss2006rigid}.}
%

The MV basis, the dual canonical basis, and the dual semicanonical basis are all examples of biperfect bases \cite{baumann2019mirkovic}.  In \cite[Appendix]{baumann2019mirkovic}, we showed that the the MV basis differs from the dual semicanonical basis.  Since the basis vector at which these bases differ is not a cluster monomial, this gives evidence for the following conjecture of Anderson and Kogan \cite[Conjecture 5.1]{anderson2006algebra}.
\begin{conjecture}\label{conj:cluster in MV}
    %The dual semicanonical basis and the MV basis overlap on the cluster monomials. 
    The cluster monomials are contained in the MV basis. 
\end{conjecture}

The conjecture is true in type $A_n$ for $n\leq 3$ 
% as the MV basis and the dual semicanonical basis agree in these cases 
since in these cases $\CC[N]$ has a unique biperfect basis (C.f.\ \cite{baumann2019mirkovic}).
% 
More generally, cluster monomials supported on seeds satisfying certain extra conditions were shown to be in the MV basis in \cite{baumann2020bases}. 
Also, cluster variables coming from the initial seeds constructed in \cite{berenstein2005cluster3} are in the MV basis \cite{baumann2019mirkovic}.

% \acom{Re the above paragraph, I like how it was: It has been shown in \cite{kato2011polytopal} that for initial seeds $(x_1, \dots, x_r)$ obtained from a reduced expression of the longest word in the Weyl group for $G$, each $x_i$ is in the MV basis as these correspond to generalized minors introduced in \cite{berenstein2005cluster3} created in \cite{geiss2007initial} for the dual semicanonical basis. If the reduced expression satisfies a special condition, then the cluster monomials in the initial seed are also in the MV basis \cite{baumann2020bases}.}

%This leads to the following idea to show that the cluster variables are in the MV basis. Given a seed $((x_1,\cdots,x_r),B)$ with each of the $x_i$ in the MV basis, say corresponding to MV cycles $Z_i^*$, and an exchange relation $x_ix_i^* = x_+ + x_-$, it suffices to find an MV cycle $Z_i^*$ such that the fusion product of $Z_i$ and $Z_i^*$ gives $Z_+$ and $Z_-$, corresponding respectively to $x_+$ and $x_-$. In this case, we are able to conclude that $x_i^*$ is in the MV basis corresponding to the MV cycle $Z_i^*$. Inducting on mutation, we conclude that all cluster variables are in the MV basis.

In particular, we hope that each exchange relation can be realized using the fusion of MV cycles. If so, we are then able to conclude that all cluster variables are in the MV basis. We show in Section \ref{ss:GL4 examples} that this is indeed the case in type $A_3$, by directly computing each fusion product corresponding to an exchange relation.

% {\color{blue}
% \acom{To quote from BFZ for a concise definition: Cluster algebras are a class of commutative rings defined axiomatically in terms of a distinguished family of generators called cluster variables. Say also what is a seed, what is a mutation, what is a cluster, what is a cluster monomial, and finally for the stubbornly ignorant clarify that 'the cluster monomials' refers to cluster monomials for all possible clusters!}

% $\CC[N]$ is a cluster algebra as shown in \cite[\S 4.3]{geiss2007initial} using \cite[Theorem~2.10]{berenstein2005cluster3}, and one of the main motivations for this paper is using fusion to check that certain cluster monomials are contained in the MV basis.
% that are associated to exchange relations in $\CC[N]$ with its cluster algebra structure.

% \acom{Background on cluster:}

% In the appendix to \cite{baumann2019mirkovic} the second and third author along with Morton-Ferguson check that the MV basis differs from the dual semicanonical basis (at the Kashiwara--Saito counterexample). 
% Yet, the two bases have enough in common (c.f.\ BKT,...?), that we have the following conjecture.

% \begin{conjecture}
%     The dual semicanonical basis and the MV basis overlap on the cluster monomials. In particular, the cluster monomials are contained in the MV basis. 
% \end{conjecture}

% We note that both of these bases are \new{biperfect} bases of $\CC[N]$ and in type $A_n$ with $n\le 3$ there is a unique biperfect basis of $\CC[N]$. }

% Anderson and Kogan conjectured in \cite{anderson2006algebra} \acom{or earlier?} that those MV polynomials which are cluster monomials for a Fomin--Zelevinsky cluster algebra structure on $\CC[N]$ are naturally expressible as determinants\dots
% % and they conjectured a formula for many of them.

% It's not clear how this work helps with/relates to AK/their conjectures.

% The generalized \mvy is interesting in its own right. \acom{is it?}

% Computing fusion still hard but at least boiled down to linear algebra. Cf.\ fusion product as it appears in \cite{beilinson1991quantization,feigin2generalized,mirkovic2007geometric,anderson2006algebra,bezrukavnikov2005equivariant}.

% Interesting combinatorics? Presumably it is related to Feigin and Loktev's generalized Kostka polynomials? Though Joel says that the product on SSYT that we are witnessing is unlikely to have a combinatorial description.

% Exchange relations only work on cluster modules where one is a mutation of the other (i.e.\ those corresponding to cluster monomials which are related by mutation). Of course this gives me everything. Up to $A_4$ as in type $A_5$ there exist indecomposable modules which are not cluster, so exchange relations do not apply.

% The hope (conjecture) is that this paper gives a way to compute on such modules. Cf.\ counterexample satisfying $\barD(c_Y) = \barD(b_Z) + 2\barD(b)$, where $b$ is a cluster monomial belonging to both bases and $c_Y$ is the square of a cluster monomial ($b'$ say) belonging to both bases. The $\barD$ equation suggests that $(b')^2 = b_Z + 2b$. We can try to check this using Roger's script. Such calculations bring us closer to answering two questions. First, how forgetful is $\barD$. Second, on the level of modules, are the terms appearing in the fusion product determined by $\Ext^1$ of the factors. By Roger's thesis this is true if the factors are cluster modules related by a mutation.

% % No; $x*y=\sum z\Rightarrow y = \frac 1 x \sum z$.

% Representation theory?

\section{Supporting Cast}
\label{s:players}
\subsection{Rings and discs}
\label{ss:rings}
Set $\cO = \cO_0 = \CC\xt$ and $\cK = \cK_0 = \CC\xT$.
% TODO: I rearranged this sentence, because starting it with 'for any other' presupposes we have introduced s 
%  For any other 
We will also consider $\cO_s = \CC\xt[t-s]$ and its fraction field $\cK_s=\CC\xT[t-s]$, for any 
$ s \in \CC\setminus\{0\} $, as well as $ \Oinf = \CC\xt[t^{-1}] $ and $\Kinf = \CC\xT[t^{-1}]$. For any point $ s \in \PP = \PP^1$, 
% TODO: Identifying $s\in\CC$ and the line $s\in\PP$ containing it right?
$ \cO_s$ is the completion of the local ring $ \mathcal O_{\PP, s} $ and thus the formal spectrum of $ \cO_s$ is the formal neighbourhood $ D_s$ of $ s $ (also called the formal disc centered at $ s$).  Similarly the formal spectrum of the field $\cK_s$ is the deleted formal neighbourhood (or punctured disc), denoted $ D_s^\times$.

Note that for $ s \in \CC $, we have an obvious isomorphisms  $\cO_s \cong \cO$ and $\cK_s\cong\cK$ taking $ t-s $ to $ t$. % ok technically the point s is in P and the coordinate s is valued in C 
% 
\subsection{Groups}
\label{ss:groups}
% 
Let $H $ be an algebraic group over $ \CC $.  We will be interested in $ H(R)$ where $ R $ is a $\CC$-algebra, for example $R = \CC[t], \cO_s$, etc.  
% 
% \acom{random question, I know $H(R)$ is supposed to mean functor of points of the scheme $H$ but can it also be viewed as a base change from $\CC$ to $\RR$?}
% \jcom{Usually $ H_R$ is the notation used for the base change, so using $ H(R)$ here is a bit of an abuse, but it is common and harmless.}
% 
Note that evaluation at $ t = s$ provides a group homomorphism $ H(\cO_s) \rightarrow H$.  We denote the kernel of this map by $ H_1(\cO_s)$, often called the  first congruence subgroup.  We will be particularly interested in this construction in the case $ s = \infty$, which gives us the group $ H_1(\Oinf)$.
% \acom{Add $G_1$}
%\acom{TODO: Algebraic groups over rings, i.e.\ the meaning of $G(K)$. The affine space of $m\times m$ matrices which we denote $ M_m$. The first congruence subgroup $G_1$ and why we take $\Oinf$ points instead of $\CC[t]$ points as some of the lit does.}

>>>>

%!TEX program = lualatex
% 
\documentclass{article} 
\usepackage{basic}
\usepackage{float}
\title{Computing fusion products of MV cycles using the Mirkovi\'c--Vybornov isomorphism}
\author{Roger Bai, Anne Dranowski, Joel Kamnitzer} 
\date{\today}

@@

\maketitle
% \tableofcontents
\begin{abstract}
    Motivated by a conjecture of Anderson describing the structure of the ring of functions on a unipotent subgroup as an algebra spanned by the Mirkovi\'c--Vilonen cycles for the Langlands dual group with respect to the fusion product, we give a conceptually elementary approach to computing this product in type $A$.

    We do so with the help of the Mirkovi\'c--Vybornov isomorphism (one of several established in \cite{mirkovic2007quiver}), which relates slices in the affine Grassmannian of $\GL_m$ and slices in nilpotent orbits of $N\times N$ matrices.

    Extending this isomoprhism to one relating families of slices in the two point Beilinson--Drinfeld Grassmannian and families of slices in orbits of matrices having two eigenvalues, we are able to transfer the fusion product of cycles to a fusion of matrices.

    As an application, we confirm that the cluster variables in $\CC[N]$ are contained in its Mirkovi\'c--Vilonen basis when $m=4$. 
\end{abstract}

\section{Introduction}
\label{s:intro}
% 
\subsection{Geometric Satake and Mirkovi\'c--Vilonen cycles}
\label{ss:gsmv}
% 
The geometric Satake equivalence of Mirkovi\'c and Vilonen \cite{mirkovic2007geometric} is an equivalence of categories between the category of spherical perverse sheaves on the affine Grassmannian of a reductive group $ G $ and the category of representations of its Langlands dual group $G^\vee$.  This foundational result provides a powerful tool to study representation theory using geometry.  In particular, under this equivalence, the MV cycles (short for Mirkovi\'c--Vilonen) in the affine Grassmannian of $ G $ index bases for irreducible representations of $ G^\vee $. In this paper we work with $ G = \GL_m$ and identify $ G^\vee = \GL_m$ as well.

In \cite{mirkovic2007quiver} (see also the recent sequel \cite{mirkovic2019comparison}) Mirkovi\'c and Vybornov supply a geometric version of symmetric and skew Howe $(\GL_m,\GL_N)$ dualities.  They relate Kazhdan--Lusztig slices in the affine Grassmannian of $\GL_m$ to slices in $N\times N$ nilpotent orbits on the one hand, and to Nakajima quiver varieties on the other hand.    The second author \cite{dthesis} showed  that under the Mirkovi\'c--Vybornov isomorphism, the MV cycles are identified with certain varieties of matrices, called generalized orbital varieties.

Essential to the geometric Satake equivalence, the BD Grassmannian (short for Beilinson--Drinfeld) is used to define a fusion product on the category of spherical perverse sheaves on the affine Grassmannian.  In \cite{anderson2003polytope}, Anderson used the BD Grassmannian to define a fusion of MV cycles.  He constructed a family where the general fibre is a Cartesian product of two MV cycles, and the special fibre, called the fusion, is a (non-reduced) union of MV cycles.  Anderson conjectured that this fusion product matches the multiplication in the coordinate ring of the unipotent subgroup $ N \subset\GL_m$.  This conjecture was proven in \cite{baumann2019mirkovic}, where it was also established that the MV cycles give a basis for $ \CC[N]$.

\subsection{Fusion using generalized orbital varieties}
\label{ss:fugov}
% 
Computing the fusion product of MV cycles using the BD Grassmannian is quite difficult, both computationally and conceptually.  A few such computations were done by Anderson and Kogan in \cite{anderson2006algebra} in a somewhat ad-hoc fashion.  Another perspective on these computations (using convolution Grassmannians) was recently studied by Baumann, Gaussent and Littelmann \cite{baumann2020bases}.

The purpose of this paper is to give a conceptually elementary way to compute this product by transferring it to a fusion product of generalized orbital varieties.  To do so, we use that the Mirkovi\'c--Vybornov isomorphism extends to an isomorphism between families of slices in the BD Grassmannian and families of slices of matrices with two eigenvalues. Our main result is the following (see \Cref{cor:goodreps} for a more precise statement).

\begin{theorem}
    Let $ Z', Z'' $ be MV cycles.  The structure constants for the multiplication $ b_{Z'} b_{Z''}$ in $\CC[N]$ are equal to the intersection multiplicities of generalized orbital varieties in the fusion of the generalized orbital varieties corresponding to $ Z', Z''$.
\end{theorem}

@@

\subsection{Cluster algebras and MV cycles}
One motivation for computing products of MV basis elements is to relate the cluster algebra structure on $\CC[N]$ to the MV basis. Cluster algebras were introduced by Fomin and Zelevinsky in \cite{fomin2002cluster1} and it was shown in \cite{geiss2007initial}, using a construction of \cite{berenstein2005cluster3}, that $\CC[N]$ has a natural cluster algebra structure.

This structure starts with an initial seed $\Sigma = (\{x_1,\dots,x_r\},B)$, where $x_i \in \CC[N]$ are a cluster of variables, $B$ is a certain matrix, and $ r = m(m-1)/2$.
We obtain new variables through a process called mutation, where for a mutable variable $x_i$, its mutation is the unique element $x_i^* \in \CC[N]$ such that 
$$
    x_ix_i^* = x_+ + x_- \,.
$$
This equation is called an exchange relation, and the terms $x_+$ and $x_-$ are certain monomials in the $x_j$, $j\neq i$,  determined by $B$. 
It gives us the new cluster $\{x_1, \dots, x_i^*, \dots, x_r\}$.

The initial cluster $\{x_1,\dots,x_r\}$ as well as all possible variables that are obtained through successive mutations are known as cluster variables. 
The cluster monomials are products of cluster variables that are supported on a single cluster.

There are two well-known bases of $\CC[N]$ whose relationship with the cluster structure has been heavily studied: the dual canonical basis of \cite{lusztig1990canonicalbases} and the dual semicanonical basis of \cite{lusztig2000semicanonical}.
In \cite{kang2018monoidal}, Kang, Kashiwara, Kim and Oh showed that the cluster monomials are contained in the dual canonical basis, while, in \cite{geiss2006rigid}, Gei\ss, Leclerc and Schr\"oer proved that they are all contained in the dual semicanonical basis.
% \cite[Theorem~2.8(3)]{geiss2006rigid}.
%

The MV basis, the dual canonical basis, and the dual semicanonical basis are all examples of biperfect bases of $\CC[N]$ \cite[Section 2]{baumann2019mirkovic}. In particular, they share a crystal structure.
In \cite[Appendix]{baumann2019mirkovic}, we showed that the MV basis differs from the dual semicanonical basis at the same point in the underlying crystal as where the dual semicanonical basis differs from the dual canonical basis.  
% 
The fact that this point at which each of these bases differ is not a cluster monomial in any of these bases, gives evidence for the following conjecture of Anderson and Kogan \cite[Conjecture 5.1]{anderson2006algebra}.
\begin{conjecture}\label{conj:cluster in MV}
    The cluster monomials are contained in the MV basis. 
\end{conjecture}

The conjecture is true in type $A_n$ for $n\leq 3$ since in these cases $\CC[N]$ has a unique biperfect basis \cite[Section 2.3]{baumann2019mirkovic}.
% 
More generally, cluster variables called flag minors, coming from initial seeds constructed in \cite[Definitions~2.2--2.3]{berenstein2005cluster3}, belong to each biperfect basis of $\CC[N]$ \cite[Remark~2.10]{baumann2019mirkovic}. 
And recently in \cite[Proposition 7.2]{baumann2020bases} cluster monomials supported on clusters of flag minors satisfying a certain combinatorial condition where shown to be in the MV basis.

Our hope is that each exchange relation can be realized using the fusion of MV cycles. This would imply that all cluster variables are in the MV basis. In \Cref{ss:GL4 examples} we verify that this is indeed the case in type $A_3$ by directly computing each fusion product corresponding to an exchange relation.

\section{Supporting Cast}
\label{s:players}
% 
\subsection{Rings and discs}
\label{ss:rings}
Set $\cO = \cO_0 = \CC\xt$ and $\cK = \cK_0 = \CC\xT$.
We will also consider $\cO_s = \CC\xt[t-s]$ and its fraction field $\cK_s=\CC\xT[t-s]$, for any $ s \in \CC\setminus\{0\} $, as well as $ \Oinf = \CC\xt[t^{-1}] $ and $\Kinf = \CC\xT[t^{-1}]$. For any point $ s \in \PP = \PP^1$, $ \cO_s$ is the completion of the local ring $ \mathcal O_{\PP, s} $ and thus the formal spectrum of $ \cO_s$ is the formal neighbourhood $ D_s$ of $ s $ (also called the formal disc centered at $ s$).  Similarly, the formal spectrum of the field $\cK_s$ is the deleted formal neighbourhood (or punctured disc), denoted $ D_s^\times$.

Note that for $ s \in \AA $, we have an obvious isomorphism  $\cO_s \cong \cO$ and $\cK_s\cong\cK$ taking $ t-s $ to $ t$. 
%

\subsection{Groups}
\label{ss:groups}
% 
Let $H $ be an algebraic group over $ \CC $.  We will be interested in $ H(R)$ where $ R $ is a $\CC$-algebra, for example $R = \CC[t], \cO_s$, etc.  
% 
Note that evaluation at $ t = s$ provides a group homomorphism $ H(\cO_s) \rightarrow H$.  We denote the kernel of this map by $ H_1(\cO_s)$, often called the  first congruence subgroup.  We will be particularly interested in this construction in the case $ s = \infty$, which gives us the group $ H_1(\Oinf)$.

Throughout the paper, we fix $ m \in \NN$.  We let $G = \GL_m$ and we let $T\subset G$ be the maximal torus of diagonal matrices. 
We identify $\ZZ^m$ with the coweight lattice of $G$ and say that a coweight $\nu = (\nu_1,\dots,\nu_m)$ is \new{dominant} if $\nu_1\ge\cdots\ge\nu_m$ and \new{effective} if $\nu_j\ge 0$ for all $j$.
% 
If $\nu$ is both effective and dominant, then it is a partition of size $|\nu| = \nu_1 + \cdots + \nu_m\in\NN$.  Let $ Q_+ \subset \ZZ^m$ denote the positive root cone, explicitly we have
$$
    Q_+ = \{ (\nu_1, \dots, \nu_m) : \nu_1 + \cdots + \nu_j \ge 0 \text{ for $j = 1, \dots, m-1 $ and } \nu_1 + \dots + \nu_m = 0 \}\,. 
$$
We define a partial order on $ \ZZ^m$ by $ \lambda \ge \mu $ if and only if $ \lambda - \mu \in Q_+ $.

Given $s\in\CC$ we define $(t-s)^\nu$ to be the diagonal matrix 
\[
\begin{bmatrix}
    (t-s)^{\nu_1} \\
    & (t-s)^{\nu_2} \\ 
    & & \ddots \\
    & & & (t-s)^{\nu_m}
\end{bmatrix} 
\]
which we can view in $G(K)$ for any ring $ K $ containing $(t-s)^{-1}$ and $\CC[t]$. For example, $(t-s)^\nu\in G(K)$ for $ K = \Ks$ or $\CC(t)$.

We will also be interested in the affine space $ M_m$ of $m\times m$ matrices.  Note that for any $ \CC$-algebra $ R $, $G(R)$ consists of those matrices $ g \in M_m(R) $ whose determinant is invertible in $ R$. Thus for example $ (t-s)^\mu\in M_m(\CC[t])$ for all effective $ \mu \in \ZZ^m$ but $(t-s)^\mu\in G(\CC[t])$ if and only if $ \mu = 0 $.
%

\subsection{Lattices}
\label{ss:lat}
% 
We will use the lattice model for the affine Grassmannian, so it is useful to recall the following definition. Let $ R \subset K$ be two $\CC$-algebras (usually, but not always, $K$ will be a field). 
Consider $ K^m $ as a $K$-module. By restriction $ K^m$ can be viewed as an $R$-module.  An \new{$R$-lattice} in $K^m$ is an $R$-submodule $ L \subset K^m$ which is a free $R$-module of rank $ m $ and satisfies $ L \otimes_R K = K^m $. Equivalently, $ L = \Sp_R(v_1, \dots, v_m)$ where $v_1, \dots, v_m$ are free generators of $K^m$. 
%

$ R^m \subset K^m $ is called the standard lattice. The group $\GL_m(K) $ acts transitively on the set of $R$-lattices in $K^m$, thus giving a bijection between this set and $\GL_m(K)/\GL_m(R)$, since $ \GL_m(R) $ is the stabilizer of the standard lattice.

We will be particularly interested in $\CC[t]$-lattices in $ \CC(t)^m$. Given such a lattice $ L $ and a point $ a \in \CC$, the \new{specialization} of $ L $ at $ a $ is the lattice in $\cK_a^m$ defined as $ L(a) := L \otimes_{\CC[t]} \cO_a $. 
If $L(a) = \cO_a^m$ then $L$ is said to be \new{trivial at $a$}. 
For example, the lattice $(t-s)^{-1}\CC[t] \subset \CC(t)$ is trivial at any $a\ne s$, since $ t-s $ is invertible in $ \cO_a$.

\section{Affine Grassmannians} 
\label{s:affgrs}
% 
We now define various versions of the affine Grassmannian which will play important roles in this paper. 
% 
Each definition is made group-theoretically and then restated as a moduli space of vector bundles and as a moduli space of lattices.
% 
We also sketch how to pass between descriptions.

In these definitions, $ \Vtriv$ denotes the trivial rank $ m $ vector bundle.

\begin{definition}
\label{def:gr}
     The \new{ordinary affine Grassmannian} $\Gr = G(\cK)/G(\cO)$.
\end{definition}    
It is the moduli space of vector bundles with trivializations,
$$
\Gr = 
    \left\{ 
        (V, \varphi) : \text{$V$ is a rank $m$ vector bundle on $ D_0 $, $\varphi : V \xrightarrow{\sim} \Vtriv $ on $ D_0^\times $} 
    \right\} \,. 
$$
It is also a moduli space of lattices, 
$$ 
\Gr = 
    \left\{ L \subset \cK^m : \text{ $L$ is a $\cO$-lattice} \right\}\,.
$$
We obtain a lattice from a pair $ (V,\varphi) $ by setting $ L = \Gamma(D_0, V)$ which is embedded into $ \cK^m = \Gamma(D_0^\times, \Vtriv)$ using $ \varphi$.  
On the other hand, to get a lattice from the group-theoretic description $ G(\cK) / G(\cO) $ we set $ L = g \cO^m$ for $ g \in G(\cK)$.
% 
\begin{definition}
\label{def:grs}
    For any $ s \in \CC $, the \new{ordinary affine Grassmannian} $\Gr_s = G(\cK_s)/G(\cO_s)$ \new{at} $ s $. 
\end{definition}  
% 
As above, we have modular and lattice descriptions:
% 
\begin{gather*}
\Gr_s = 
    \left\{ (
        V, \varphi) : \text{$V$ is a rank $m$ vector bundle on $ D_s $, $\varphi : V \xrightarrow{\sim} \Vtriv $ on $ D^\times_s $} 
    \right\}\,, \\
\Gr_s = 
    \left\{ 
        L \subset \cK_s^m : \text{ $L$ is a $\cO_s$-lattice} 
    \right\} \,. 
\end{gather*}
% 
\begin{definition}
\label{def:grth}
    The \new{thick affine Grassmannian} $\Grth = G(\Kinf)/G(\CC[t])$.
\end{definition}
Again, we have the following modular and lattice descriptions:
\begin{gather*}
\Grth = 
    \left\{ 
        (V, \varphi) : \text{$V$ is a rank $m$ vector bundle on $ \PP $, $\varphi : V \xrightarrow{\sim} \Vtriv $ on $ D_\infty $ } 
    \right\}\,, \\
\Grth = 
    \left\{ 
        L : L \subset  \Kinf^m \text{ is a $\CC[t]$-lattice} 
    \right\}\,.
\end{gather*}
% 
\begin{definition} 
\label{def:bdgr}
    The two point \new{Beilinson--Drinfeld Grassmannian} $$\pi : \Grbd\to \AA$$ with one point fixed at 0, and the second point $s\in\AA$ varying.
\end{definition}
% 
It is described in modular terms by 
$$
\Grbd = 
    \left\{ 
        (V,\varphi,s) : V\text{ is a rank $m$ vector bundle on }\PP, \varphi : V \xrightarrow{\sim} \Vtriv \text{ on } \PP \setminus \{0, s\}  
    \right\}\,. 
$$
% 
The fibre of $\Grbd \to \AA$ over $ s \in \AA $ will be denoted $ \Gr_{0,s} $ and is given by
% 
$$ 
    G(\CC[t, t^{-1}, (t-s)^{-1}]/G(\CC[t])\,. % := \pi^{-1}(s) 
$$
We also have the lattice descriptions:
% 
\begin{gather*}
\Grbd = 
    \left\{ 
        (L,s) : L \subset  \CC(t)^m\text{ is a $\CC[t]$-lattice trivial at any }a \ne 0, s 
    \right\}\,, \\
% , $L$ is 
\Gr_{0,s} = 
    \left\{ 
        L : L \subset  \CC[t,t^{-1},(t-s)^{-1}]^m \text{ is a $\CC[t]$-lattice} 
    \right\}\,.
\end{gather*}
% 
\begin{definition} 
\label{def:grplus}
The \new{positive part} of $\Gr $, resp.\ $\Grth$, is defined by 
\begin{gather*}
    \Gr^+ = \left(M_m(\cO) \cap G(\cK)\right) / G(\cO)\,,\,\, 
    \text{resp.\ } {\Grth}^+ = \left(M_m(\CC[t]) \cap G(\Kinf)\right) / G(\CC[t])\,. 
\end{gather*}
\end{definition}
% 
In modular terms, $\Gr^+$ (resp.\ ${\Grth}^+$) is the set of those $ (V, \varphi)$ where $ \varphi : V \rightarrow \Vtriv $ extends to an inclusion of coherent sheaves over $ D_0 $ (resp.\ over $ \PP$).

In lattice terms, $ \Gr^+$ (resp.\ ${\Grth}^+$) contains those lattices $L$ which are contained in $\cO^m$ (resp.\ $\CC[t]^m$).
%
